Toy Transit Surveys

### 2017-08-31 memos/workdir/toy_analytic_surveys.tex
memo: fix typo
memo: add analytic mag-limited prob(q)

While a closed analytic expression for $\langle N_d \rangle$ does exist, it is 
messy, and it does not yield much intuition. Instead, summarizing the important 
points and supporting them numerically:
\begin{itemize}
	\item In a volume-limited sample of binary star systems in which the 
	primary mass is fixed, and the mass ratio is drawn from a bounded uniform 
	distribution, the distribution of $\gamma_R$ will be biased towards low 
	values ($\gamma_R \approx 0.1$). This is shown in 
	Fig.~\ref{fig:gammaR_distribn_vol_limited}.
	\item In a magnitude-limited sample of binary star systems in which the 
	primary mass is fixed, and the mass ratio of the \textit{population} is 
	drawn from a bounded uniform distribution, the observed distribution of 
	mass ratios will be biased towards high values. You will see more twins,
	because they are detectable out to a greater distance\footnote{This is a 
	scarcy systematic w.r.t. the claimed intrinsic excess of twin binaries.}. 
	This is shown in 
	Fig.~\ref{fig:q_distribn_mag_limited}.
	\item In a magnitude-limited sample of binary star systems in which the 
	primary mass is fixed, the distribution of $\gamma_R$ will be biased 
	towards low values ($\approx 0.1$), but less so than in a 
	volume-limited sample. This is shown in 
	Fig.~\ref{fig:gammaR_distribn_mag_limited}.
\end{itemize}

@@

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/gammaR_distribn_vol_limited.pdf}
	\end{center}
	\caption{The distribution of the luminosity ratio for a volume limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation from 
		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio}.}
	\label{fig:gammaR_distribn_vol_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/q_distribn_mag_limited.pdf}
	\end{center}
	\caption{The distribution of the mass ratio for a magnitude limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation from 
		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
		volume-limited sample!}
	\label{fig:q_distribn_mag_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/gammaR_distribn_mag_limited.pdf}
	\end{center}
	\caption{The distribution of the luminosity ratio for a magnitude limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation from 
		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio}.}
	\label{fig:gammaR_distribn_mag_limited}
\end{figure}

@@

\section{Model \# 3: A Synthetic {\it Kepler} Analog}

>>>>

While a closed analytic expression for $\langle N_d \rangle$ does exist, it is 
messy, and it does not yield much intuition. Instead, summarizing the important 
points and supporting them numerically:
\begin{itemize}
	\item In a volume-limited sample of binary star systems in which the 
	primary mass is fixed, and the mass ratio is drawn from a bounded uniform 
	distribution, the distribution of $\gamma_R$ will be biased towards low 
	values ($\gamma_R \approx 0.1$). This is shown in 
	Fig.~\ref{fig:gammaR_distribn_vol_limited}.
	\item In a magnitude-limited sample of binary star systems in which the 
	primary mass is fixed, and the mass ratio of the \textit{population} is 
	drawn from a bounded uniform distribution, the observed distribution of 
	mass ratios will be biased towards high values. You will see more twins,
	because they are detectable out to a greater distance\footnote{This is a 
	scary systematic w.r.t. the claimed intrinsic excess of twin binaries.}. 
	This is shown in Fig.~\ref{fig:q_distribn_mag_limited}, and explained 
	analytically below.
	\item In a magnitude-limited sample of binary star systems in which the 
	primary mass is fixed, the distribution of $\gamma_R$ will be biased 
	towards low values ($\approx 0.1$), but less so than in a 
	volume-limited sample. This is shown in 
	Fig.~\ref{fig:gammaR_distribn_mag_limited}.
\end{itemize}

@@

Similarly, the probability distribution function for the mass ratio of binaries 
in a magnitude limited sample can be found by marginalizing over the joint 
distribution for a binary star's position and mass ratio.
Explicitly, using `ml' as a subscript for `magnitude-limited', and assuming 
$L\propto M^3$,
\begin{align}
{\rm prob}_{\rm ml}(q) &= 
	\int_{0}^{d_{\rm max}(q)} {\rm prob}_{\rm ml}(r,q) \,{\rm d}r \\
&= \mathcal{Z}\cdot (1+q^3)^{3/2}\quad {\rm if\ }0.1<q\leq 1,
\label{eq:pdf_q_mag_limited}
\end{align}
and otherwise 0, for $\mathcal{Z}$ the appropriate normalization constant.
Eq.~\ref{eq:pdf_q_mag_limited} is the ``analytic'' form overplotted in 
Fig.~\ref{fig:q_distribn_mag_limited}. Its independence from both the primary 
star mass as well as from the magnitude cutoff will be used in 
Sec.~\ref{sec:model_3}.
The form given in Eq.~\ref{eq:pdf_q_mag_limited} is found by writing 
${\rm prob}_{\rm ml}(r,q) = {\rm prob}(q|r){\rm prob}_{\rm ml}(r)$, using 
Eq.~\ref{eq:mass_ratio} for ${\rm prob}(q|r)$, and marginalizing over the mass 
ratio
\begin{align}
{\rm prob}_{\rm ml}(r) &= \int_{0}^{1} {\rm prob}_{\rm ml}(r|q)\,{\rm 
prob}(q)\,{\rm d}q
\\
&= \int_{0.1}^{1} \frac{1}{3} \frac{r^2}{d_{\rm max}^3(q)}\,{\rm d}q
\quad {\rm if\ }r<d_{\rm max}(q),
\end{align}
and otherwise zero. The resulting joint probability distribution is plotted in 
Fig.~\ref{fig:joint_prob_r_q}.

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/gammaR_distribn_vol_limited.pdf}
	\end{center}
	\caption{The distribution of the luminosity ratio for a volume limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation from 
		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio}.}
	\label{fig:gammaR_distribn_vol_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/q_distribn_mag_limited.pdf}
	\end{center}
	\caption{The distribution of the mass ratio for a magnitude limited 
		sample of binary stars. For numerics, ${\rm BF} = 0.45$ [Duchene and 
		Kraus 2013], total number density from Bovy 2017, empirical $M(L)$ 
		from Fig.~\ref{fig:mass_luminosity}, and mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
		volume-limited sample!
		The analytic line (Eq.~\ref{eq:pdf_q_mag_limited}) is shown to indicate 
		that the entire effect, albeit with a simpler $M(L)$, can be understood 
		without resorting to numerics. 
		}
	\label{fig:q_distribn_mag_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/gammaR_distribn_mag_limited.pdf}
	\end{center}
	\caption{The distribution of the luminosity ratio for a magnitude limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation from 
		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio}.}
	\label{fig:gammaR_distribn_mag_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/joint_prob_r_q.pdf}
	\end{center}
	\caption{The joint probability distribution of a binary star's position and 
	mass ratio in a magnitude-limited sample. This plot assumes a different 
	$M(L)$ relation than previous plots: $L=M^3$. Note that the marginalized 
	mass ratio distribution is uniform at any given distance.}
	\label{fig:joint_prob_r_q}
\end{figure}

@@

\section{Model \# 3: A Synthetic {\it Kepler} Analog}
\label{sec:model_3}
